Carbonate chemistry at Ny-Ålesund

### 2021-08-02 awipev-CO2.Rmd
clean-up and fiddle with date discrete data

```{r read and prepare discrete AT+CT, echo=FALSE, warning=FALSE, message=FALSE}
d <- tibble(
  read.table(paste0(path, "fb_data/Discrete_analyses_AT_CT/Discrete_sampling_AWIPEV.csv"), header = T, dec = ".", as.is = T, sep = ";", fill = TRUE)
) %>% 
  dplyr::mutate(datetime = dmy_hm(sampling_date, tz="UTC"),
                measure_date = dmy(measure_date, tz="UTC"),
                date = as.Date(datetime)) %>% 
  dplyr::rename(ph_s_sf = pH_s_seafet,
                ph_s_dur = pH_s_durafet,
                qflag_ph_s_sf = qflag_pH_s_seafet,
                qflag_ph_s_dur = qflag_pH_s_durafet)

@@

# Conversion of spec pH of in situ samples to in situ temp and 4 °C
pH_seafet <- d %>%
  dplyr::filter(!is.na(ph_s_sf) & !is.na(temp_insitu_ctd))
pH_seafet <- pH_seafet %>% 
  dplyr::mutate(
    ph_s_sf_temp_insi = pHinsi(pH = pH_seafet$ph_s_sf, 
                         ALK = pH_seafet$at_mean_interp*1e-6, Tinsi = pH_seafet$temp_insitu_ctd, 
                         Tlab = pH_seafet$temp_s_lab, Pinsi = pH_seafet$depth/10, 
                         S = pH_seafet$sal_mix, Pt = 0, Sit = 0, k1k2 = "l", kf = "dg", ks="d", 
                         pHscale = "T", b="u74"),
   ph_s_sf_temp_4 = pHinsi(pH = pH_seafet$ph_s_sf, 
                         ALK = pH_seafet$at_mean_interp*1e-6, Tinsi = 4, 
                         Tlab = pH_seafet$temp_s_lab, Pinsi = pH_seafet$depth/10, 
                         S = pH_seafet$sal_mix, Pt = 0, Sit = 0, k1k2 = "l", kf = "dg", ks="d", 
                         pHscale = "T", b="u74")
  )
# Conversion of spec pH of FerryBox samples to FerryBox temp
pH_durafet <- d %>%
  tidyr::drop_na(ph_s_dur, sal_fb, temp_fb)
pH_durafet <- pH_durafet %>% 
  dplyr::mutate(
    ph_s_dur_temp_fb = pHinsi(pH = pH_durafet$ph_s_dur, 
                              ALK = pH_durafet$at_mean_interp*1e-6, Tinsi = pH_durafet$temp_fb, 
                              Tlab = pH_durafet$temp_s_lab, Pinsi = 0, S = pH_durafet$sal_fb, 
                              Pt = 0, Sit = 0, k1k2 = "l", kf = "dg", ks="d", pHscale = "T", b="u74"),
   ph_s_dur_temp_4_l = pHinsi(pH = pH_durafet$ph_s_dur, # lueker
                         ALK = pH_durafet$at_mean_interp*1e-6, Tinsi = 4, 
                         Tlab = pH_durafet$temp_s_lab, Pinsi = 0, 
                         S = pH_durafet$sal_fb, Pt = 0, Sit = 0, k1k2 = "l", kf = "dg", ks="d", 
                         pHscale = "T", b="u74"),
   ph_s_dur_temp_4_s20 = pHinsi(pH = pH_durafet$ph_s_dur, # Sulpis
                         ALK = pH_durafet$at_mean_interp*1e-6, Tinsi = 4, 
                         Tlab = pH_durafet$temp_s_lab, Pinsi = 0, 
                         S = pH_durafet$sal_fb, Pt = 0, Sit = 0, k1k2 = "s20", kf = "dg", ks="d", 
                         pHscale = "T", b="u74"),       # Millero 02
   ph_s_dur_temp_4_p18 = pHinsi(pH = pH_durafet$ph_s_dur, # Papadimitriou
                         ALK = pH_durafet$at_mean_interp*1e-6, Tinsi = 4, 
                         Tlab = pH_durafet$temp_s_lab, Pinsi = 0, 
                         S = pH_durafet$sal_fb, Pt = 0, Sit = 0, k1k2 = "p18", kf = "dg", ks="d", 
                         pHscale = "T", b="u74"),
   ph_s_dur_temp_4_m02 = pHinsi(pH = pH_durafet$ph_s_dur,  # Millero 02
                         ALK = pH_durafet$at_mean_interp*1e-6, Tinsi = 4, 
                         Tlab = pH_durafet$temp_s_lab, Pinsi = 0, 
                         S = pH_durafet$sal_fb, Pt = 0, Sit = 0, k1k2 = "m02", kf = "dg", ks="d", 
                         pHscale = "T", b="u74")

   )
# Make mean for replicates + SD
pH_seafet_mean <- pH_seafet %>%
  dplyr::select(date, ph_s_sf_temp_insi, ph_s_sf_temp_4) %>% 
  dplyr::group_by(date) %>% 
  dplyr::summarise(ph_s_sf_temp_insi_mean = mean(ph_s_sf_temp_insi, na.rm = TRUE),
                   #ph_s_sf_temp_insi_sd = sd(ph_s_sf_temp_insi, na.rm = TRUE),
                   ph_s_sf_temp_4_mean = mean(ph_s_sf_temp_4, na.rm = TRUE)
                   )
pH_durafet_mean <- pH_durafet %>%
  dplyr::select(date, ph_s_dur_temp_fb, ph_s_dur_temp_4_l, ph_s_dur_temp_4_s20,
                ph_s_dur_temp_4_p18, ph_s_dur_temp_4_m02) %>% 
  dplyr::group_by(date) %>% 
  dplyr::summarize(ph_s_dur_temp_fb_mean = mean(ph_s_dur_temp_fb, na.rm = TRUE),
                   #ph_s_dur_temp_fb_sd = sd(ph_s_dur_temp_fb, na.rm = TRUE),
                   ph_s_dur_temp_4_l_mean = mean(ph_s_dur_temp_4_l, na.rm = TRUE),
                   ph_s_dur_temp_4_s20_mean = mean(ph_s_dur_temp_4_s20, na.rm = TRUE),
                   ph_s_dur_temp_4_p18_mean = mean(ph_s_dur_temp_4_p18, na.rm = TRUE),
                   ph_s_dur_temp_4_m02_mean = mean(ph_s_dur_temp_4_m02, na.rm = TRUE)
                   )

#  make means in "d" to have same format and call it "discrete"
discrete <- d %>%
  dplyr::group_by(date) %>%
  dplyr::summarize(
    datetime = mean(datetime, na.rm = TRUE),
    sal_fb = mean(sal_fb, na.rm = TRUE),
    sal_mix = mean(sal_mix, na.rm = TRUE),
    temp_fb = mean(temp_fb, na.rm = TRUE),
    at_mean_interp = mean(at_mean_interp, na.rm = TRUE),
    at = mean(at, na.rm = TRUE),
    #qflag_at = mean(qflag_at, na.rm = TRUE), # already dealt with above
    ct_mean_interp = mean(ct_mean_interp, na.rm = TRUE),
    ct = mean(ct, na.rm = TRUE),
    #qflag_ct = mean(qflag_ct, na.rm = TRUE)  # already dealt with above
    temp_mix = mean(temp_mix, na.rm = TRUE),
    temp_insitu_11m = mean(temp_insitu_11m, na.rm = TRUE)
  )
# join discrete + seafet/durafet
discrete <- left_join(discrete, pH_seafet_mean, by='date')
discrete <- left_join(discrete, pH_durafet_mean, by='date')

@@

# create new variable in "discrete_is_na" to bind "discrete_no_na"
discrete_is_na$ph_calc_fb <- NA
discrete_is_na$pco2_calc_fb <- NA
discrete <- bind_rows(discrete_no_na, discrete_is_na) %>% 
  dplyr::arrange(date)
```
```{r read and prepare nutrient data, echo=FALSE, warning=TRUE, message=FALSE}
# File "Ferrybox nutrients 2014-2018.xlsx" provided by 
# Willem van de Poll <w.h.van.de.poll@rug.nl> on 2020-01-16
# nutrients in umol/l
nuts <- readxl::read_excel(paste0(path, "fb_data/Ferrybox nutrients 2014-2018.xlsx"),
                           skip =1) %>% 
  drop_na(date) %>% 
  dplyr::mutate(datetime = as.POSIXct(paste0(date, " 12:00:00")),
                ) %>% # assuming sampling time 12:00
  dplyr::filter(datetime > as.POSIXct("2015-07-09"))
# The nearest times
time_index <- unique(d_all$datetime)
# nuts$closest <- time_index[
#   FNN::knnx.index(time_index, # The values to be queried (big df)
#                   as.vector(nuts$datetime), # The value to find the closest match to (small df)
#                   k = 1)] # The number of closest values to return
nuts <- nuts %>% 
  dplyr::mutate(closest = time_index[
  FNN::knnx.index(time_index, # The values to be queried (big df)
                  as.vector(nuts$datetime), # The value to find the closest match to (small df)
                  k = 1)]) %>%  # The number of closest values to return
  dplyr::select(-date, -datetime) %>% 
  dplyr::rename(NO3NO2 = `NO3+NO2`)
# cleanup and convert in umol/kg
d_all <- left_join(d_all, nuts, by = c("datetime" = "closest")) #%>% 
  #dplyr::select(-closest_datetime)
d_all <- d_all %>% 
  dplyr::mutate(density_fb = seacarb::rho(S = sal_fb, T = temp_fb, P = 0),
                po4 = PO4 / (density_fb/1000),
                nh4 = NH4 / (density_fb/1000),
                no3no2 = NO3NO2 / (density_fb/1000),
                no3 = NO3 / (density_fb/1000),
                si = Si / (density_fb/1000)
  )
```

@@

# durafet pH 
discrete2merge <- discrete %>%
  dplyr::select(closest_datetime, ph_s_dur_temp_fb_mean, ph_s_dur_temp_4_l_mean,
                ph_s_dur_temp_4_s20_mean, ph_s_dur_temp_4_p18_mean, ph_s_dur_temp_4_m02_mean) %>%
  dplyr::rename(ph_s_dur_temp_fb = ph_s_dur_temp_fb_mean,
                ph_s_dur_temp_4 = ph_s_dur_temp_4_l_mean, # lueker by default
                ph_s_dur_temp_4_l = ph_s_dur_temp_4_l_mean,
                ph_s_dur_temp_4_s20 = ph_s_dur_temp_4_s20_mean,
                ph_s_dur_temp_4_p18 = ph_s_dur_temp_4_p18_mean,
                ph_s_dur_temp_4_m02 = ph_s_dur_temp_4_m02_mean,) %>%
  dplyr::filter(!is.na(closest_datetime)) %>%
  rename(datetime=closest_datetime)
d_all <- full_join(d_all, discrete2merge, by="datetime")

@@

# In order to remove the temperature effect from the observed pCO2 , the observed pCO2 values are normalized to a constant temperature (annual mean) -- Takahashi et al. (2002)
temp_mean <- mean(taka$temp_insitu_11m, na.rm = TRUE)
taka <- taka %>% 
  dplyr::mutate(pco2_mean_temp = pco2_fb_insi_temp * exp(0.0423 * (temp_mean - temp_insitu_11m)))
d_pco2_bio <- max(taka$pco2_mean_temp, na.rm = TRUE) - min(taka$pco2_mean_temp, na.rm = TRUE)
d_pco2_temp <- max(taka$pco2_fb_insi_temp, na.rm = TRUE) - min(taka$pco2_fb_insi_temp, na.rm = TRUE)
temp_bio <- d_pco2_temp / d_pco2_bio
```

>>>>

```{r read and prepare discrete AT+CT, echo=FALSE, warning=FALSE, message=FALSE}
d <- tibble(
  read.table(paste0(path, "fb_data/Discrete_analyses_AT_CT/Discrete_sampling_AWIPEV.csv"), header = T, dec = ".", as.is = T, sep = ";", fill = TRUE)
) %>% 
  dplyr::mutate(datetime = ymd_hm(sampling_date, tz="UTC"),
                measure_date = ymd(measure_date, tz="UTC"),
                date = as.Date(datetime)) %>% 
#  dplyr::filter(datetime < as.POSIXct("2021-01-01 00:00:00")) %>% #eliminate 2021 data
  dplyr::rename(ph_s_sf = pH_s_seafet,
                ph_s_dur = pH_s_durafet,
                qflag_ph_s_sf = qflag_pH_s_seafet,
                qflag_ph_s_dur = qflag_pH_s_durafet)

@@

# Conversion of spec pH of in situ samples to in situ temp and 4 °C
pH_seafet <- d %>%
  dplyr::filter(!is.na(ph_s_sf) & !is.na(temp_insitu_ctd))
pH_seafet <- pH_seafet %>% 
  dplyr::mutate(
    ph_s_sf_temp_insi = pHinsi(pH = pH_seafet$ph_s_sf, 
                         ALK = pH_seafet$at_mean_interp*1e-6, Tinsi = pH_seafet$temp_insitu_ctd, 
                         Tlab = pH_seafet$temp_s_lab, Pinsi = pH_seafet$depth/10, 
                         S = pH_seafet$sal_mix, Pt = 0, Sit = 0, k1k2 = "l", kf = "dg", ks="d", 
                         pHscale = "T", b="u74"),
   ph_s_sf_temp_4 = pHinsi(pH = pH_seafet$ph_s_sf, 
                         ALK = pH_seafet$at_mean_interp*1e-6, Tinsi = 4, 
                         Tlab = pH_seafet$temp_s_lab, Pinsi = pH_seafet$depth/10, 
                         S = pH_seafet$sal_mix, Pt = 0, Sit = 0, k1k2 = "l", kf = "dg", ks="d", 
                         pHscale = "T", b="u74")
  )
# Conversion of spec pH of FerryBox samples to FerryBox temp
pH_durafet <- d %>%
  tidyr::drop_na(ph_s_dur, sal_fb, temp_fb)
pH_durafet <- pH_durafet %>% 
  dplyr::mutate(
    ph_s_dur_temp_fb = pHinsi(pH = pH_durafet$ph_s_dur, 
                              ALK = pH_durafet$at_mean_interp*1e-6, Tinsi = pH_durafet$temp_fb, 
                              Tlab = pH_durafet$temp_s_lab, Pinsi = 0, S = pH_durafet$sal_fb, 
                              Pt = 0, Sit = 0, k1k2 = "l", kf = "dg", ks="d", pHscale = "T", b="u74"),
   ph_s_dur_temp_4 = pHinsi(pH = pH_durafet$ph_s_dur, # lueker
                         ALK = pH_durafet$at_mean_interp*1e-6, Tinsi = 4, 
                         Tlab = pH_durafet$temp_s_lab, Pinsi = 0, 
                         S = pH_durafet$sal_fb, Pt = 0, Sit = 0, k1k2 = "l", kf = "dg", ks="d", 
                         pHscale = "T", b="u74"),

   # ph_s_dur_temp_4_l = pHinsi(pH = pH_durafet$ph_s_dur, # lueker
   #                       ALK = pH_durafet$at_mean_interp*1e-6, Tinsi = 4, 
   #                       Tlab = pH_durafet$temp_s_lab, Pinsi = 0, 
   #                       S = pH_durafet$sal_fb, Pt = 0, Sit = 0, k1k2 = "l", kf = "dg", ks="d", 
   #                       pHscale = "T", b="u74"),
   # ph_s_dur_temp_4_s20 = pHinsi(pH = pH_durafet$ph_s_dur, # Sulpis
   #                       ALK = pH_durafet$at_mean_interp*1e-6, Tinsi = 4, 
   #                       Tlab = pH_durafet$temp_s_lab, Pinsi = 0, 
   #                       S = pH_durafet$sal_fb, Pt = 0, Sit = 0, k1k2 = "s20", kf = "dg", ks="d", 
   #                       pHscale = "T", b="u74"),       # Millero 02
   # ph_s_dur_temp_4_p18 = pHinsi(pH = pH_durafet$ph_s_dur, # Papadimitriou
   #                       ALK = pH_durafet$at_mean_interp*1e-6, Tinsi = 4, 
   #                       Tlab = pH_durafet$temp_s_lab, Pinsi = 0, 
   #                       S = pH_durafet$sal_fb, Pt = 0, Sit = 0, k1k2 = "p18", kf = "dg", ks="d", 
   #                       pHscale = "T", b="u74"),
   # ph_s_dur_temp_4_m02 = pHinsi(pH = pH_durafet$ph_s_dur,  # Millero 02
   #                       ALK = pH_durafet$at_mean_interp*1e-6, Tinsi = 4, 
   #                       Tlab = pH_durafet$temp_s_lab, Pinsi = 0, 
   #                       S = pH_durafet$sal_fb, Pt = 0, Sit = 0, k1k2 = "m02", kf = "dg", ks="d", 
   #                       pHscale = "T", b="u74")

   )
# Make mean for replicates + SD
pH_seafet_mean <- pH_seafet %>%
  dplyr::select(date, ph_s_sf_temp_insi, ph_s_sf_temp_4) %>% 
  dplyr::group_by(date) %>% 
  dplyr::summarise(ph_s_sf_temp_insi_mean = mean(ph_s_sf_temp_insi, na.rm = TRUE),
                   ph_s_sf_temp_insi_sd = sd(ph_s_sf_temp_insi, na.rm = TRUE),
                   ph_s_sf_temp_4_mean = mean(ph_s_sf_temp_4, na.rm = TRUE)
                   )
pH_durafet_mean <- pH_durafet %>%
  dplyr::select(date, ph_s_dur_temp_fb, ph_s_dur_temp_4) %>% 
  dplyr::group_by(date) %>% 
  dplyr::summarize(ph_s_dur_temp_fb_mean = mean(ph_s_dur_temp_fb, na.rm = TRUE),
                   ph_s_dur_temp_fb_sd = sd(ph_s_dur_temp_fb, na.rm = TRUE),
                   ph_s_dur_temp_4_mean = sd(ph_s_dur_temp_4, na.rm = TRUE))

#bind all spectro measurements, just to assess sd, change name of dataframe only to do the rbind.
seafetph <- pH_seafet_mean
colnames(seafetph) <- c("date", "mean", "sd")
durafetph <- pH_durafet_mean
colnames(durafetph) <- c("date", "mean", "sd")
pH_seafet_durafet <- rbind(seafetph,durafetph)
#calculations for all pH spectro measurement for sd
sd_min <- min(pH_seafet_durafet$sd)
sd_max <- max(pH_seafet_durafet$sd)
sd_mean <- mean(pH_seafet_durafet$sd)
# =======
#                    #ph_s_dur_temp_fb_sd = sd(ph_s_dur_temp_fb, na.rm = TRUE),
#                    ph_s_dur_temp_4_l_mean = mean(ph_s_dur_temp_4_l, na.rm = TRUE),
#                    ph_s_dur_temp_4_s20_mean = mean(ph_s_dur_temp_4_s20, na.rm = TRUE),
#                    ph_s_dur_temp_4_p18_mean = mean(ph_s_dur_temp_4_p18, na.rm = TRUE),
#                    ph_s_dur_temp_4_m02_mean = mean(ph_s_dur_temp_4_m02, na.rm = TRUE)
#                    )
# >>>>>>> 0d3b58ee3abb01c904f165290b5c76bfe7270c1f

#put back the right column name for the following script
pH_durafet_mean$ph_s_dur_temp_fb <- pH_durafet_mean$ph_s_dur_temp_fb_mean
pH_seafet_mean$ph_s_sf_temp_insi <- pH_seafet_mean$ph_s_sf_temp_insi_mean
#  make means in "d" to have same format and call it "discrete"
discrete <- d %>%
  dplyr::group_by(date) %>%
  dplyr::summarize(
    datetime = mean(datetime, na.rm = TRUE),
    sal_fb = mean(sal_fb, na.rm = TRUE),
    sal_mix = mean(sal_mix, na.rm = TRUE),
    temp_fb = mean(temp_fb, na.rm = TRUE),
    at_mean_interp = mean(at_mean_interp, na.rm = TRUE),
    at = mean(at, na.rm = TRUE),
    #qflag_at = mean(qflag_at, na.rm = TRUE), # already dealt with above
    ct_mean_interp = mean(ct_mean_interp, na.rm = TRUE),
    ct = mean(ct, na.rm = TRUE),
    #qflag_ct = mean(qflag_ct, na.rm = TRUE)  # already dealt with above
    temp_mix = mean(temp_mix, na.rm = TRUE),
    temp_insitu_11m = mean(temp_insitu_11m, na.rm = TRUE)
  )
# join discrete + seafet/durafet
discrete <- left_join(discrete, pH_seafet_mean, by='date')
discrete <- left_join(discrete, pH_durafet_mean, by='date')

@@

# create new variable in "discrete_is_na" to bind "discrete_no_na"
discrete_is_na$ph_calc_fb <- NA
discrete_is_na$pco2_calc_fb <- NA
discrete <- bind_rows(discrete_no_na, discrete_is_na) %>% 
  dplyr::arrange(date)
```

```{r read and prepare nutrient data, echo=FALSE, warning=TRUE, message=FALSE}
# File "Ferrybox nutrients 2014-2018.xlsx" provided by 
# Willem van de Poll <w.h.van.de.poll@rug.nl> on 2020-01-16
# nutrients in umol/l
nuts <- readxl::read_excel(paste0(path, "fb_data/Ferrybox nutrients 2014-2018.xlsx"),
                           skip =1) %>% 
  drop_na(date) %>% 
  dplyr::mutate(datetime = as.POSIXct(paste0(date, " 12:00:00")),
                ) %>% # assuming sampling time 12:00
  dplyr::filter(datetime > as.POSIXct("2015-07-09"))
# The nearest times
time_index <- unique(d_all$datetime)
# nuts$closest <- time_index[
#   FNN::knnx.index(time_index, # The values to be queried (big df)
#                   as.vector(nuts$datetime), # The value to find the closest match to (small df)
#                   k = 1)] # The number of closest values to return
nuts <- nuts %>% 
  dplyr::mutate(closest = time_index[
  FNN::knnx.index(time_index, # The values to be queried (big df)
                  as.vector(nuts$datetime), # The value to find the closest match to (small df)
                  k = 1)]) %>%  # The number of closest values to return
  dplyr::select(-date, -datetime) %>% 
  dplyr::rename(NO3NO2 = `NO3+NO2`)
# cleanup and convert in umol/kg
d_all <- left_join(d_all, nuts, by = c("datetime" = "closest")) #%>% 
  #dplyr::select(-closest_datetime)
d_all <- d_all %>% 
  dplyr::mutate(density_fb = seacarb::rho(S = sal_fb, T = temp_fb, P = 0),
                po4 = PO4 / (density_fb/1000),
                nh4 = NH4 / (density_fb/1000),
                no3no2 = NO3NO2 / (density_fb/1000),
                no3 = NO3 / (density_fb/1000),
                si = Si / (density_fb/1000)
  )
```

@@

# durafet pH 
discrete2merge <- discrete %>%
  dplyr::select(closest_datetime, ph_s_dur_temp_fb_mean, ph_s_dur_temp_4_mean) %>%
  dplyr::rename(ph_s_dur_temp_fb = ph_s_dur_temp_fb_mean,
                ph_s_dur_temp_4 = ph_s_dur_temp_4_mean # lueker by default
                ) %>%
  dplyr::filter(!is.na(closest_datetime)) %>%
  rename(datetime=closest_datetime)
d_all <- full_join(d_all, discrete2merge, by="datetime")

@@

# In order to remove the temperature effect from the observed pCO2, the observed pCO2 values are normalized to a constant temperature (annual mean) -- Takahashi et al. (2002)
temp_mean <- mean(taka$temp_insitu_11m, na.rm = TRUE)
taka <- taka %>% 
  dplyr::mutate(pco2_mean_temp = pco2_fb_insi_temp * exp(0.0423 * (temp_mean - temp_insitu_11m)))
d_pco2_bio <- max(taka$pco2_mean_temp, na.rm = TRUE) - min(taka$pco2_mean_temp, na.rm = TRUE)
d_pco2_temp <- max(taka$pco2_fb_insi_temp, na.rm = TRUE) - min(taka$pco2_fb_insi_temp, na.rm = TRUE)
temp_bio <- d_pco2_temp / d_pco2_bio
```
